---
title: "practice"
author: "Jordan Moody"
date: "4/10/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mdsr)
library(RMySQL)
db <- dbConnect_scidb(dbname = "imdb")
```

## Making table of number of movies in each rating
```{r}
db %>%
  dbGetQuery("SELECT *
             FROM info_type;")

```
```{r}
db %>%
  dbGetQuery("SELECT *
             FROM movie_info
             WHERE info_type_id = 97
             LIMIT 0,10;")
```
```{r}
db %>%
  dbGetQuery("SELECT *
             FROM title
             LIMIT 0, 10")
```
### Gross is info_type_id = 107
Country is info_type_id = 8 (filter for USA)
MPAA is info_type_id = 97 (filter '%Rated R%' or similar)
```{r}
db %>%
  dbGetQuery("SELECT t.production_year, sum(m.info) as N, 
             FROM title t
             JOIN movie_info m ON m.movie_id = t.id
             WHERE t.production_year > 1967
             AND t.info_type = 1
             AND m.info_type_id = 97
             AND m.info = '%Rated R%'
             GROUP BY t.production_year;")

```
###Make table displaying the number of each rating per year since 1968


##Steps:
Make table of number of films with each rating for each year???
Other things
Join with table of grosses to look at grosses for each movie??? (Possibly use sums)